\documentclass{article}

\usepackage{geometry} 
\geometry{letterpaper} 

\title{Instructions for Generating SSC Documentation}
\author{Paul Gilman}

\begin{document}

\maketitle

\begin{abstract}
This document explains how to use Doxygen, LaTeX, and HeVeA to create the SSC API documentation.
\end{abstract}


\section{Overview}

Generating the SSC documentation involves a number of steps, some of which are automated, and some of which are automatic.

\begin{itemize}
\item The SSC documentation consists of the 8 chapters in the \texttt{ssc\_guide.tex} file.

\item The content for Chapters 1-7 is in \texttt{ssc\_guide.tex}. It was originally drafted by Aron Dobos and revised by Paul Gilman.

\item Doxygen generates content for Chapter 8 (the function reference for \texttt{sscapi.h}) in \texttt{sscapi\_8h.tex} from comments in \texttt{sscapi.h}.

\item An include command in the \texttt{ssc\_guide.tex} file inserts the content of \texttt{sscapi\_8h.tex} into \texttt{ssc\_guide.tex}.

\item The settings in the Doxygen configuration file \texttt{ssc\_api\_doxyconfig} determine what is included in the function reference: Only a single page with the function reference, LaTeX document rather than HTML, etc.

\item TeXworks generates the final PDF document from LaTeX.

\item HeVeA converts the PDF document to HTML.

\item Manually modify the CSS in the HTML document to format for the web.

\end{itemize}

\section{Overall Steps for Maintaining Documentation}

\begin{enumerate}

\item As needed, revise content in \texttt{ssc\_guide.tex} and Doxygen-formatted comments in \texttt{sscapi.h}.

\item Run Doxygen to generate the LaTeX document from Doxygen-formatted comments in the \texttt{sscapi.h} header file.

\item Run TeXworks to convert LaTex document to PDF. (Run it twice to generate TOC, cross references, etc.)

\item \textbf{Skip this if not generating HTML:} Run HeVeA to convert PDF to HTML.

\end{enumerate}

\section{Notes}

\begin{itemize}

\item Run Doxygen on \texttt{ssc\_api\_doxyconfig}. Doxygen creates the \texttt{ssc\_api\_latex} folder.

\item Ignore \texttt{doxygen.sty} that doxygen creates in the in \texttt{ssc\_api\_latex} folder, and use the custom \texttt{doxygen.sty} in the SVN repository instead. It removes numbering of individual functions, and a few other minor formatting changes.

\textbf{Note.} If you use the \texttt{doxygen.sty} file that doxygen automatically generates instead of the customized one, open \texttt{doxygen.sty} in a text editor, and comment out or delete the commands under "Setup fancy headings" so that the "Generated on date for SSC API by Doxygen" footer does not appear on pages.

\item Run TexWorks on \texttt{ssc\_guide.tex} to generate PDF. (Run twice to update table of contents.)

\item \textbf{Skip this if not generating HTML:} Run HeVeA on \texttt{ssc\_guide.pdf} to generate HTML: hevea \texttt{ssc\_guide.tex}.

\end{itemize}
\section{Software Installation Instructions}


\subsection{LaTeX tools}

\subsection{Doxygen}

\subsection{HeVeA}

\end{document}